\documentclass{report}

\title{Solutions Manual for Lawvere and Schanuel \emph{Conceptual Mathematics}}
\author{Robin Adams}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pf2}

\newcommand{\inv}[1]{\ensuremath{{#1}^{-1}}}

\begin{document}
    \maketitle
    \tableofcontents

    \part{Preview}
    \chapter{Session 1 --- Galileo and Multiplication of Objects}
    \paragraph{Exercise 1}
    Many examples --- every instance of a product in a category gives an example. I will not list them.

    \paragraph{Exercise 2}
    I am not entirely sure what solution the authors had in mind. Here are some that come to my mind:

    Place a spirit level between the two points and see if it reads as level.

    Place a smooth plank between the two points and see if a ball placed at one point rolls to the other,
    or \emph{vice versa}.

    Hang a plumbline at each point and see if they form a right angle with the line joining the two points.

    Of these, the third is my favourite.

    \part{Part I --- The category of sets}

    \chapter{Article I --- Sets, maps, composition}

    \paragraph{Exercise 1}
    Easy.

    \paragraph{Exercise 2}
    There are 8 maps from $A$ to $B$.

    \paragraph{Exercise 3}
    There are 27 maps from $A$ to $A$.

    \paragraph{Exercise 4}
    There are 9 maps from $B$ to $A$.

    \paragraph{Exercise 5}
    There are 4 maps from $B$ to $B$.

    \paragraph{Exercise 6}
    There are 10 such maps from $A$ to $A$.

    \paragraph{Exercise 7}
    There are 3 such maps from $B$ to $B$.

    \paragraph{Exercise 8}
    There is no such pair of maps.

    \paragraph{Exercise 9}
    There are 12 such pairs of maps.

    \chapter{Session 3 --- Composing maps and counting maps}

    \paragraph{Exercise 1}
    (a) and (c) make sense.

    \paragraph{Exercise 2}
    (a) and (c) still make sense.

    \part{Part II --- The algebra of composition}

    \chapter{Article II --- Isomorphisms}

    \section{1. Isomorphisms}

    \paragraph{Exercise 1}
    \subparagraph{(R)}
    We have $1_A \circ 1_A = 1_A$ by the Identity Laws, so $1_A$ is an isomorphism with inverse $1_A$.

    \subparagraph{(S)}
    We have $g \circ f = 1_A$ and $f \circ g = 1_B$ (this is what it means for $g$ to be an inverse for $f$).
    This says exactly that $f$ is an inverse for $g$.

    \subparagraph{(T)}
    Let $\inv{f} : B \rightarrow A$ be an inverse for $f$ and $\inv{k} : C \rightarrow B$ be an inverse for $k$.
    We prove $\inv{f} \circ \inv{k}$ is an inverse for $k \circ f$. We have
    \begin{align*}
        \inv{f} \circ \inv{k} \circ k \circ f & = \inv{f} \circ 1_B \circ f & (\text{definition of inverse}) \\
        & = \inv{f} \circ f & (\text{Identity Law})\\
        & = 1_A & (\text{definition of inverse})
    \end{align*}
    and $k \circ f \circ \inv{f} \circ \inv{k} = 1_C$ similarly.

    \paragraph{Exercise 2}
    We have
    \begin{align*}
        g & = g \circ 1_B & (\text{Identity Law}) \\
        & = g \circ f \circ k & (\text{$k$ is an inverse of $f$}) \\
        & = 1_A \circ k & (\text{$g$ is an inverse of $f$}) \\
        & = k & (\text{Identity Law})
    \end{align*}

    \paragraph{Exercise 3}
    \subparagraph{(a)}
    Let $f : A \rightarrow B$. Let $h, k : C \rightarrow A$.

    Suppose $f \circ h = f \circ k$. Then
    \begin{align*}
        \inv{f} \circ f \circ h & = \inv{f} \circ f \circ k \\
        \therefore 1_A \circ h & = 1_A \circ k & (\text{Definition of inverse}) \\
        \therefore h & = k & (\text{Identity Law}) \qed
    \end{align*}

    \subparagraph{(b)}
    Let $f : A \rightarrow B$. Let $h, k : B \rightarrow C$.

    Suppose $h \circ f = k \circ f$. Then
    \begin{align*}
        h \circ f \circ \inv{f} & = k \circ f \circ \inv{f} \\
        \therefore h \circ 1_B & = k \circ 1_B & (\text{Definition of inverse}) \\
        \therefore h & = k & (\text{Identity Law}) \qed
    \end{align*}

    \subparagraph{(c)}
    Let $A = \{ 0,1 \}$. Define $f : A \rightarrow A$ by $f(0) = 1$ and $f(1) = 0$. Define $h : A \rightarrow A$
    by $h(x) = 0$ for all $x$. Define $k : A \rightarrow A$ by $k(x) = 1$ for all $x$.

    $f$ is invertible, and is its own inverse.

    We have $h \circ f = f \circ k = h$.

    We do not have $h = k$.

    \paragraph{Exercise 4}
    \subparagraph{(1)}
    This function is invertible with inverse $\inv{f}(x) = (x-7)/3$.

    \subparagraph{(2)}
    This function is invertible with inverse $\inv{g}(x) = \sqrt{x}$.

    \subparagraph{(3)}
    This function is not invertible because $h(1) = h(-1) = 1$.

    \subparagraph{(4)}
    This function is not invertible because $k(1) = k(-1) = 1$.

    \subparagraph{(5)}
    This function is not invertible because there is no $x$ such that $l(x) = 2$.

    \section{2 --- General division problems: Determination and choice}

    \paragraph{Exercise 5}
    There are 6 maps $f$ such that $g \circ f = 1_{\{0,1\}}$; we can map 0 to any of $b$, $p$ or $q$,
    and 1 to either of $r$ or $s$.

    Given any one of these maps $f$, there are 8 maps $g$ such that $g \circ f = 1_{\{0,1\}}$.
    We must map $f(0)$ to $0$, $f(1)$ to $1$, and the other three elements to any of 0 or 1.

    \paragraph{Exercise 6}
    If $r : B \rightarrow A$ is a section for $f$, then we take $t = g \circ r$. We have
    $t \circ f = g \circ r \circ f = g \circ 1_A = g$.

    \paragraph{Exercise 7}
    Let $s : B \rightarrow A$ be a section for $f$. Let $T$ be any set and $t_1, t_2 : T \rightarrow B$.
    Suppose $t_1 \circ f = t_2 \circ f$. Then
    \begin{align*}
        t_1 \circ f \circ s & = t_2 \circ f \circ s \\
        \therefore t_1 \circ 1_B & = t_2 \circ 1_B \\
        \therefore t_1 & = t_2
    \end{align*}

    \paragraph{Exercise 8}
    If $s_1 : B \rightarrow A$ is a section for $r_1 : A \rightarrow B$ and $s_2 : C \rightarrow B$
    is a section for $r_2 : B \rightarrow C$, then $s_1 \circ s_2$ is a section for $r_2 \circ r_1$ since
    \begin{align*}
        r_2 \circ r_1 \circ s_1 \circ s_2 & = r_2 \circ 1_B \circ s_2 \\
        & = r_2 \circ s_2 \\
        & = 1_C
    \end{align*}

    \paragraph{Exercise 9}
    We have
    \begin{align*}
        e \circ e & = f \circ r \circ f \circ r \\
        & = f \circ 1 \circ r & (\text{$r$ is a retraction of $f$}) \\
        & = f \circ r \\
        & = e
    \end{align*}

    \paragraph{Exercise 10}
    From the proof of Proposition 3, $\inv{f} \circ \inv{g}$ is both a section and a retraction for
    $g \circ f$.

    \paragraph{Exercise 11}
    Set $f(Fatima) = coffee$, $f(Omer) = tea$ and $f(Alysia) = cocoa$. Then $f$ is an isomorphism.

    There is no isomorphism $g : A \rightarrow C$. For if $g(Fatima) = true$ then $g(Omer)$ must be $false$,
    and then it is impossible to choose a value for $g(Alysia)$ without having $g(Alysia) = g(Fatima)$
    or $g(Alysia) = g(Omer)$. Similarly if $g(Fatima) = false$ then $g(Omer)$ must be $true$, and then
    again we cannot choose a value for $g(Alysia)$.

    \chapter{Session 4 --- Division of Maps: Isomorphisms}

    \section{4. A small zoo of isomorphisms in other categories}

    \paragraph{Exercise 1}
    We have $h(d(x)) = h(2x) = x$ and $d(h(x)) = d(x/2) = x$ for any $x$.

    \paragraph{Exercise 2}
    $f(odd) = negative$ and $f(even) = positive$

    \paragraph{Exercise 3}
    \subparagraph{(a)}
    This is not an isomorphism because $p(0 + 0) = 1$ but $p(0) + p(0) = 2$

    \subparagraph{(b)}
    This is not an isomorphism because it is not surjective; there is no $x$ such that $sq(x) = -1$.

    \subparagraph{(c)}
    This is not an isomorphism because it is not injective. We have $sq(1) = sq(-1) = 1$.

    \subparagraph{(d)}
    This is an isomorphism; it is bijective and $-(x+y) = (-x) + (-y)$.

    \subparagraph{(e)}
    This is not an isomorphism because $m(1 \times 1) = -1$ but $m(1) \times m(1) = 1$.

    \subparagraph{(f)}
    This is not a well-defined map because $c(-1) = -1 \notin \mathbb{R}_{>0}$.

    \chapter{Session 5 --- Division of Maps: Sections and Retractions}

    \section{1. Determination Problems}

    \paragraph{Exercise 1}
    \subparagraph{(a)}
    Suppose $h = g \circ f$ and $f a_1 = f a_2$. Then $h a_1 = g(f a_1) = g(f a_2) = h a_2$.

    \subparagraph{(b)}
    No. Take $A = C = \emptyset$ and $B = \{ * \}$. Let $f : A \rightarrow B$ and $h : A \rightarrow C$
    be the unique such maps. Vacuously, if $f a_1 = f a_2$ then $h a_1 = h a_2$. But there is no map
    $g : B \rightarrow C$.

    \section{3. Choice Problems}

    \paragraph{Exercise 2}
    \subparagraph{(a)}
    Suppose $g \circ f = h$. Let $a \in A$. Let $b = f(a)$. Then $h(a) = g(f(a)) = g(b)$.

    \subparagraph{(b)}
    This is equivalent to the Axiom of Choice.

    \section{5. Stacking or Sorting}

    \paragraph{Exercise 3}
    I'm not going to draw all of them, but there are 8 of them.

    \chapter{Session 9 --- Retracts and Idempotents}

    \section{1. Retracts and Comparisons}

    \paragraph{Exercise 1}
    If $A$ is empty, then the nowhere-defined function is a map $A \rightarrow B$.

    If $B$ has a point, say $b$, then the constant map with value $b$ is a map $A \rightarrow B$.

    \section{2. Idempotents as records of retracts}

    \paragraph{Exercise 3}
    Suppose $s : A \rightarrow B$, $r : B \rightarrow A$ and $s' : A' \rightarrow B$, $r' : B \rightarrow A'$
    are splittings of $e : B \rightarrow B$. Let
    \begin{align*}
        f & = r' \circ s & : A \rightarrow A' \\
        \inv{f} & = r \circ s' & : A' \rightarrow A
    \end{align*}
    Then we have
    \begin{align*}
        f \circ \inv{f} & = r' \circ s \circ r \circ s' \\
        & = r' \circ e \circ s' \\
        & = r' \circ s' \circ r' \circ s' \\
        & = 1 \\
        \inv{f} \circ f & = r \circ s' \circ r' \circ s \\
        & = r \circ e \circ s \\
        & = r \circ s \circ r \circ s \\
        & = 1
    \end{align*}

    \chapter{Quiz}

    \paragraph{Question 1}
    Let $A = \{ * \}$ and $B = \{ 0,1 \}$. Define $f : A \rightarrow B$ by $f(*) = 0$. Then the unique
    function $r : B \rightarrow A$ is a retraction for $f$ (since $r(f(*)) = *$) but not a section for $f$
    (since $f(s(1)) = 0$). Therefore there is no section for $f$, since there is only one map $B \rightarrow A$.

    \paragraph{Question 2}
    \subparagraph{(a)}
    Yes: if $pqp = p$ then $pqpq = pq$
    \subparagraph{(b)}
    Yes: if $pqp = p$ then $qpqp = qp$

    \paragraph{Question 2*}
    Let $q' = qpq$ Then we have
    \begin{align*}
        pq'p & = pqpqp \\
        & = pqp \\
        & = p \\
        q'pq' & = qpqpqpq \\
        & = qpqpq \\
        & = qpq \\
        & = q'
    \end{align*}

    \paragraph{Question 1*}
    Take $A = B = \mathbb{N}$ and define $f : A \rightarrow B$ by $f(x) = 2x$. Then $f$ has a retraction
    $r$ given by
    \[ r(y) = \begin{cases}
        y/2 & \text{if $y$ is even} \\
        0 & \text{if $y$ is odd}
    \end{cases} \]
    It has no section since it is not surjective (Article II Proposition 1).

    \chapter{Summary / quiz on pairs of 'opposed' maps}

    \paragraph{Question 1}
    Given two maps $f$, $g$ with domains and codomains as above, we can always form the composites
    $g \circ f$ and $f \circ g$. All we can say about $g \circ f$ and $f \circ g$ as maps in themselves is
    that they are endomaps.

    \paragraph{Question 2}
    If we know that $g$ is a retraction for $f$, that means $g \circ f$ is actually the identity map $1_A$;
    then we can prove that $f \circ g$ is not only an endomap, but actually an idempotent. The latter means
    that the equation $f \circ g \circ f \circ g = f \circ g$ is true.

    \paragraph{Question 3}
    If we even know that $f$ is an isomorphism \emph{and} that $g \circ f = 1_A$, then $f \circ g$ is not
    only an idempotent, but is the identity map $1_B$. If, moreover, $s$ is a map for which $f \circ s =
    1_B$, we can conclude that $s = g$.

    \paragraph{Question 4}
    Going back to 0, i.e. assuming no equations, but only the domain and codomain statements about $f$ and $g$,
    the composite $f \circ g \circ f$ could be different from $f$. Likewise $f \circ g \circ f \circ g$
    could be different from $f \circ g$.

    \chapter{Test 1}

    \paragraph{Question 1}
    \subparagraph{(a)}
    Let $f(Mara) = Aurelio$, $f(Aurelio) = Mara$ and $f(Andrea) = Andrea$.

    \subparagraph{(b)}
    Let $e(Mara) = Aurelio$, $e(Aurelio) = Aurelio$ and $e(Andrea) = Andrea$.

    \subparagraph{(c)}
    Let $B = \{ 0, 1 \}$. Define $s : B \rightarrow A$ by $s(0) = Aurelio$ and $s(1) = Andrea$.
    Define $r : A \rightarrow B$ by $r(Mara) = 0$, $r(Aurelio) = 0$ and $r(Andrea) = 1$.

    \paragraph{Question 2}
    Define $g : \mathbb{R} \rightarrow \mathbb{R}$ by $g(y) = (y+7)/4$.
    \subparagraph{(a)}
    $g(f(x)) = g(4x-7) = (4x-7+7)/4 = 4x/4 = x$
    \subparagraph{(b)}
    $f(g(x)) = f((x+7)/4) = 4((x+7)/4) - 7 = x+7-7 = x$

    \chapter{Session 10 --- Brouwer's Theorems}

    \section{4. Relation between fixed point and retraction theorems}

    \paragraph{Exercise 1}
    Suppose for a contradiction there is no point $x$ such that $f(x) = g(x)$. Define $r : D \rightarrow C$
    as follows: for $x \in D$, $r(x)$ is the point on $C$ that is pointed at by the arrow with tail at
    $f(x)$ and head at $g(x)$. For $x \in C$, we have $g(j(x)) = j(x)$, so the point that is pointed at by
    any arrow with head at $g(j(x))$ is $x$. Hence
    \begin{align*}
        r(j(x)) & = x
    \end{align*}
    and so $r$ is a retraction for $j$, contradicting the retraction theorem.

    \paragraph{Exercise 2}
    Let $f : A \rightarrow A$ be any endomap. Then $s \circ f \circ r : X \rightarrow X$ is an endomap on $X$.
    Hence there exists $x : T \rightarrow X$ such that $sfrx = x$. But then we have
    \begin{align*}
        rsfrx & = rx \\
        \therefore frx & = rx
    \end{align*}
    and so $r \circ x : T \rightarrow A$ is a fixed point of $f$.

    \paragraph{Exercise 3}
    Let $A$ be either $E$, $C$ or $S$, and $X$ be $I$, $D$ or $B$ respectively. Assume that every endomap $X \rightarrow X$ has a fixed point. 

    Assume for a contradiction that $X$ is a retract of $A$. By Exercise 2, every endomap
    on $A$ has a fixed point. This is a contradiction, as the antipodal map on $A$ has no fixed point.

    \section{7. Using maps to formulate guesses}

    \paragraph{Exercise 1}
    \subparagraph{(a)}
    We can express 'I start in Buffalo and end in Rochester' as $m \circ j = i \circ j$.

    We can express 'You start and finish anywhere between Buffalo and Rochester' as: there exists $f : I
    \rightarrow E$ such that $y \circ j = i \circ f$.

    \subparagraph{(b)}
    There exists $t : 1 \rightarrow I$ such that $mt = yt$.

    \subparagraph{(c)}
    Let $C$ be the circle, $D$ the disk and $P$ the plane. Let $j : C \rightarrow D$ and $i : D \rightarrow P$
    be the inclusions.

    For any maps $m, y : D \rightarrow P$ such that:
    \begin{itemize}
        \item $mj = ij$
        \item there exists $f : C \rightarrow D$ such that $yj = if$
    \end{itemize}
    then there exists $t : 1 \rightarrow D$ such that $mt = yt$.

    \subparagraph{(d)}
    I have not been able to find any smooth maps for which it is not true.

    \part{Part III --- Categories of Structured Sets}

    \chapter{Article III --- Examples of Categories}

    \section{1. The category $\mathcal{S}^\circ$ of endomaps of sets}

    \paragraph{Exercise 1}
    Let $f : (X, \alpha) \rightarrow (Y, \beta)$ and $g : (Y, \beta) \rightarrow (Z, \gamma)$. Then
    \[ g \circ f \circ \alpha = g \circ \beta \circ f = \gamma \circ g \circ f \]
    and so $g \circ f : (X, \alpha) \rightarrow (Z, \gamma)$.

    \section{4. Categories of endomaps}

    \paragraph{Exercise 2}
    Suppose $e : A \rightarrow A$ is idempotent and has a retraction $r : A \rightarrow A$. Then
    \[ 1_A = r \circ e = r \circ e \circ e = 1_A \circ e = e \]
    so $e = 1_A$. Thus, the identities are the only idempotents that have retractions.

    \paragraph{Exercise 3}
    Suppose $A$ has an even number of elements, say $\{ a_1, a_2, \ldots, a_{2n} \}$. Define $\theta : A
    \rightarrow A$ by $\theta(a_{2k+1}) = a_{2k+2}$ and $\theta(a_{2k+2}) = a_{2k+1}$ ($0 \leq k < n$). Then
    $\theta$ is an involution with no fixed point.

    Conversely, suppose $\theta : A \rightarrow A$ is an involution with no fixed point. Enumerate the elements
    of $A$ as follows: Pick any element $a_1 \in A$. Let $a_2 = \theta(a_1)$; then $a_1 = \theta(a_2)$.

    Assuming we have picked $a_1$, \ldots, $a_{2m}$ such that $\{ a_1, \ldots, a_{2m} \}$ is closed under
    $\theta$ and $A \neq \{ a_1, \ldots, a_{2m} \}$, pick $a_{2m+1} \in A - \{ a_1, \ldots, a_{2m} \}$.
    Then $\theta(a_{2m+1}) \notin \{ a_1, \ldots, a_{2m} \}$ (since $\theta(\theta(a_{2m+1})) = a_{2m+1}$)
    and $\theta(a_{2m+1}) \neq a_{2m+1}$ (since $\theta$ has no fixed point). So let
    $a_{2m+2} = \theta(a_{2m+1})$.

    This process must end because $A$ is finite. So $A = \{a_1, \ldots, a_{2n} \}$ for some $n$.

    Suppose now $A$ has an odd number of elements, say $A = \{ a_1, a_2, \ldots, a_{2n+1} \}$. Define $\theta :
    A \rightarrow A$ by
    \begin{align*}
        \theta(a_{2k+1}) & = a_{2k+2} & (0 \leq k < n) \\
        \theta(a_{2k+2}) & = a_{2k+1} & (0 \leq k < n) \\
        \theta(a_{2n+1}) & = a_{2n+1}
    \end{align*}
    Then $\theta$ is an involution whose only fixed point is $a_{2n+1}$.

    Conversely, suppose $\theta : A \rightarrow A$ is an involution with one fixed point $f$. Then
    $\theta \restriction (A - \{f\})$ is an involution on $A - \{ f \}$ with no fixed point. So $A - \{f\}$
    has an even number of elements, and so $A$ has an odd number of elements.

    \paragraph{Exercise 4}
    The map $\alpha$ is an involution because $-(-x) = x$. It is not idempotent because $-(-1) \neq -1$. Its
    only fixed point is 0.

    \paragraph{Exercise 5}
    The map $\alpha$ is not an involution because $||-1|| = 1 \neq -1$. It is idempotent because $||x|| = |x|$.
    Its fixed points are the non-negative integers.

    \paragraph{Exercise 6}
    The map $\alpha$ is an automorphism with inverse $\inv{\alpha}(x) = x - 3$.

    \paragraph{Exercise 7}
    The map $\alpha$ is not an automorphism because there is no integer $x$ with $\alpha(x) = 1$.

    \paragraph{Exercise 8}
    If $\alpha$ is idempotent then $\alpha \circ \alpha \circ \alpha = \alpha \circ \alpha = \alpha$.

    If $\alpha$ is an involution then $\alpha \circ \alpha \circ \alpha = 1 \circ \alpha = \alpha$.

    \paragraph{Exercise 9}
    Label the elements in the diagram 0, 1, 2 from top to bottom. Then
    \begin{align*}
        \alpha^3(0) & = \alpha^2(1) = \alpha(2) = 1 \\
        & = \alpha(0) \\
        \alpha^3(1) & = \alpha^2(2) = \alpha(1) = 2 \\
        & = \alpha(1) \\
        \alpha^3(2) & = \alpha^2(1) = \alpha(2) = 1 \\
        & = \alpha(2)
    \end{align*}

    Thus, $\alpha^3 = \alpha$.

    However, $\alpha$ is not idempotent because $\alpha^2(0) = 2 \neq \alpha(0)$. And $\alpha$ is not an involution
    because $\alpha^2(0) = 2 \neq 0$.

    \section{5. Irreflexive graphs}

    \paragraph{Exercise 10}
    \[ s(a) = k, s(b) = m, s(c) = k, s(d) = p, s(e) = m \]
    \[ t(a) = m, t(b) = m, t(c) = m, t(d) = q, t(e) = r \]

    The arrow $b$ has $s(b) = t(b)$. There is no arrow $x$ with $t(x) = k$.

    \paragraph{Exercise 11}
    We have
    \[ s'' \circ g \circ f = g \circ s' \circ f = g \circ f \circ s \]
    \[ t'' \circ g \circ f = g \circ t' \circ f = g \circ f \circ t \]
    and so $g \circ f : (X,P,s,t) \rightarrow (Z,R,s'',t'')$.

    \section{6. Endomaps as special graphs}
    
    \paragraph{Exericse 12}
    \[ I(g \circ f) = (g \circ f, g \circ f) = (g,g) \circ (f,f) = I(g) \circ I(f) \]

    \paragraph{Exercise 13}
    For any $x \in X$ we have $f_A(x) = 1_Y(f_A(x)) = f_D(1_X(x)) = f_D(x)$, and so $f_A = f_D$.
    Thus $(f_A, f_D) = I(f_A)$.

    \section{7. The simpler category $\mathcal{S}^\downarrow$: Objects are just maps of sets}

    \paragraph{Exercise 14}
    Let $X = \{ * \}$ and $Y = \{ 0, 1 \}$. Let $\alpha$ be the only map $X \rightarrow X$, and
    $\beta : Y \rightarrow Y$ be the map with $\beta(0) = 1$ and $\beta(1) = 0$. Let $f_A(*) = 0$
    and $f_D(*) = 1$. Then $f_D \circ \alpha = \beta \circ f_A$ but $f_A \neq f_D$.

    \section{8. Reflexive graphs}

    \paragraph{Exercise 15}
    Let $x_1 = s$ and $x_2 = t$, so $e_i = i x_i$ for each $i$. Then
    \begin{align*}
        e_k e_j & = i x_k i x_j \\
        & = i 1_P x_j \\
        & = i x_j \\
        & = e_j
    \end{align*}
    In particular, $e_j e_j = e_j$, so each $e_j$ is idempotent.

    \paragraph{Exercise 16}
    Let $(f_A, f_D) : (X, P, s, t, i) \rightarrow (Y, Q, s', t', j)$. Then
    \begin{align*}
        f_D s & = s' f_A \\
        \therefore f_D & = f_D s i \\
        & = s' f_A i
    \end{align*}

    \paragraph{Exercise 17}
    A map between $(M, F, \phi, \phi', \mu, \mu')$ and $(N, G, \psi, \psi', \nu, \nu')$
    is a pair of functions $f : M \rightarrow N$ and $g : F \rightarrow G$ such that
    \begin{align*}
        \psi f & = f \phi \\
        \psi' g & = f \phi' \\
        \nu g & = g \mu \\
        \nu' f & = g \nu
    \end{align*}

    \section{10. Retractions and Injectivity}

    \paragraph{Exercise 18}
    Let $a : X \rightarrow Y$ have a retraction $r : Y \rightarrow X$. Let $x_1, x_2 : T \rightarrow X$
    satisfy $a x_1 = a x_2$. Then
    \[ x_1 = r a x_1 = r a x_2 = x_2 \enspace . \]

    \paragraph{Exercise 19}
    We have
    \begin{align*}
        \beta a x & = 0 & a \alpha x & = 0 \\
        \beta a 0 & = 0 & a \alpha 0 & = 0
    \end{align*}
    So $\beta a = a \alpha$ as required.

    \paragraph{Exercise 20}
    Let $x_1, x_2 : (T, \gamma) \rightarrow (X, \alpha)$ satisfy $a x_1 = a x_2$. Then, for any $t \in T$,
    we have $a x_1 t = a x_2 t$, hence $x_1 t = x_2 t$ (since $a$ is injective as a function). Thus $x_1 = x_2$.

    \paragraph{Exercise 21}
    The retractions are the maps that send $y$ to $x$, $0$ to $0$, and $\overline{y}$ to either $x$ or $0$.

    \paragraph{Exercise 22}
    Let $r : Y \rightarrow X$ be either of the retractions of $a$ in $\mathcal{S}$. Then, no matter what
    $r \overline{y}$ is, we have $\alpha r \overline{y} = 0$. But $r \beta \overline{y} = r y = x$.
    Thus $r$ is not a map $(Y, \beta) \rightarrow (X, \alpha)$ in $\mathcal{S}^\circ$.

    \paragraph{Exercise 23}
    The following are maps in $\mathcal{S}^\circ$:
    \begin{align*}
        \overline{y} \mapsto x & y \mapsto 0 & 0 \mapsto 0 \\
        \overline{y} \mapsto 0 & y \mapsto 0 & 0 \mapsto 0
    \end{align*}

    \paragraph{Exericse 24}
    Suppose $(r,s) : (Y, \beta) \rightarrow (X, \alpha)$ is a retraction of $(a,a)$ in
    $\mathcal{S}^\downarrow$. Then $ra = sa = 1_X$ and $\alpha r = s \beta$. Now, whatever $r \overline{y}$ is, we have
    $\alpha r \overline{y} = 0$. But $s \beta \overline{y} = s y = x$. This is a contradiction.

    \paragraph{Exercise 25}
    Suppose $f_D s = f_D t$. For $x \in X$, we have
    \[ s' f_A x = f_D s x = f_D t x = t' f_A x \]
    and so $f_A x$ is a loop in $Y$.

    \paragraph{Exercise 26}
    Let $f : \mathbb{Z} \rightarrow \mathbb{Q}$ be the inclusion (i.e. $f(n) = n$ for every integer $n$). Then
    \subparagraph{1}
    $f$ is a map in $\mathcal{S}^\circ$ because $5f(n) = f(5n) = 5n$
    \subparagraph{2}
    The map $5 \times ()$ is an automorphism with inverse $() / 5$
    \subparagraph{3}
    If $f(m) = f(n)$ then immediately $m = n$.

    \paragraph{Exercise 27}
    Let $f : (X, \alpha) \rightarrow (Y, \beta)$. Let the two elements of $X$ be 0 and 1, where $\alpha 0 =
    \alpha 1 = 1$. Then
    \begin{align*}
        f \alpha 0 & = f \alpha 1 \\
        \therefore \beta f 0 & = \beta f 1 \\
        \therefore f 0 & = f 1
    \end{align*}
    since $\beta$ is injective. Thus $f$ is not injective.

    \paragraph{Exercise 28}
    Let $x, y \in X$ and assume $\alpha x = \alpha y$. Then
    \begin{align*}
        f \alpha x & = f \alpha y \\
        \therefore \beta f x & = \beta f y \\
        \therefore f x & = f y & (\text{$\beta$ is injective}) \\
        \therefore x & = y & (\text{$f$ is injective})
    \end{align*}

    \section{11. Types of structure}

    \paragraph{Exercise 29}
    Let $f : X \rightarrow Y$ in $\mathcal{X}$. Let $\overline{X}$ and $\overline{Y}$ be the
    $\mathcal{A}$-structures determined by $X$ and $Y$ as described in the paragraph before the exercise.
    Define $\overline{f} : \overline{X} \rightarrow \overline{Y}$ as follows. For each object $A \in
    \mathcal{A}$, we define
    \[ \overline{f}_A : \{ \text{maps} A \rightarrow X \} \rightarrow \{ \text{maps} A \rightarrow Y \} \]
    by
    \[ \overline{f}_A(g) = f \circ g \enspace . \]
    Now given a map $\alpha : A \rightarrow B$ in $\mathcal{A}$, we must prove
    $\overline{Y}_{\alpha^*} \circ \overline{f}_B =
    \overline{f}_A \circ \overline{X}_{\alpha^*}$. Well, given a map $g : B \rightarrow X$, we have
    \[ \overline{Y}_{\alpha^*}(\overline{f}_B(g)) = \overline{f}_A(\overline{X}_{\alpha^*}(g)) = f \circ g \circ \alpha \]

    \paragraph{Exercise 30}
    For every dot $a : 1 \rightarrow X$, let $e_a : S \rightarrow X$ be the map $a \circ ! : S \rightarrow 1
    \rightarrow X$ (the constant map $a$ on $S$). Then $e_a$ is an edge in the graph of $X$-fields
    and $e_a \circ s = e_a \circ t = a$, i.e. $a$ is both the source and target of $e_a$.

    If $f : X \rightarrow Y$ in $\mathcal{C}$ then the induced map sends $e_a$ to
    \[ f \circ e_a = f \circ a \circ ! = e_{fa} \]
    as required.

    \chapter{Session 11. Ascending to categories of richer structures}

    \section{1. A category of richer structures: Endomaps of sets}

    \paragraph{Exercise 1}
    There are four: we can map all three elements to the one loop, or map them to the three elements that are
    in a cycle of 3 in the second set in three different ways.

    \section{3. The category of graphs}

    \paragraph{Exercise 2}
    There are three:
    \begin{align*}
        f(a) & = p & f(b) & = r & f(c) & = q \\
        g(a) & = q & g(b) & = p & g(c) & = r \\
        h(a) & = r & h(b) & = p & h(c) & = q
    \end{align*}

    \paragraph{Exercise 3}
    Suppose for a contradiction $f : (X, \alpha) \rightarrow (Y, \beta)$. Let $x \in X$ be one of the
    elements such that $\alpha^3(x) = x$. Then $\beta^3(f(x)) = f(x)$, but there is no such element in $Y$.

    \paragraph{Exercise 4}
    For $y \in B$, we have
    \begin{align*}
        f(\alpha(\inv{f}(y))) & = \beta(f(\inv{f}(y))) \\
        & = \beta(y) \\
        \therefore \alpha(\inv{f}(y)) & = \inv{f}(\beta(y))
    \end{align*}

    \paragraph{Exercise 5}
    They are not isomorphic. In $(\mathbb{Z}, \beta)$, the elements 0, 1 and 2 are three elements such that
    none of them can be obtained by repeatedly applying $\beta$ to one of the others (they are in separate
    orbits). There are not three such elements in $(\mathbb{Z}, \alpha)$.

    \paragraph{Exercise 6}
    (a) and (d) are isomorphic; (b) and (e) are isomorphic, (c) and (f) are isomorphic.

    \paragraph{Exercise 7}
    Yes, they are isomorphic. Map the top element in the left graph to the bottom element in the right graph;
    the left element in the left graph to the left element in the right graph;
    the centre element in the left graph to the right element in the right graph;
    the right element in the left graph to the centre element in the right graph;
    the bottom element in the left graph to the top element in the right graph.

    \paragraph{Exercise 8}
    \subparagraph{(a)}
    Any path from $b$ to $e$ would be mapped by $f$ to a path from 0 to 1, but there is no such path.
    \subparagraph{(b)}
    Define $f : G \rightarrow J$ as follows. For any dot $a$ in $G$, if there is a path from $b$ to $a$
    then $f(a) = 0$; otherwise $f(a) = 1$. Map any edge $e$ from $a$ to $a'$ in $G$ to the unique edge
    from $f(a)$ to $f(a')$ in $J$. (There must be such an edge; otherwise $f(a) = 0$ and $f(a') = 1$,
    but then there is a path from $b$ to $a$ hence a path from $b$ to $a'$.)

    We have $f(b) = 0$ and $f(e) = 1$.

    \chapter{Session 12. Categories of diagrams}

    \section{1. Dynamical systems or automata}

    \paragraph{Exercise 1}
    We have $f(x') = f(\alpha^3(x)) = \beta^3(f(x)) = \beta^3(y) = y'$.

    \paragraph{Exercise 2}
    $y$ immediately enters a cycle of 3, and $z$ enters a cycle of 1 after 3 time units.

    The diagram of the light bulb has a chain of 8 different states followed by a cycle of 1.

    \section{2. Family Trees}

    \paragraph{Exercise 3}
    \subparagraph{(a)}
    For any $x \in P$, we have $gender(m(x)) = female$ and $m(gender(x)) = female$ (because $m(y) = female$
    for both $y \in G$). Likewise $gender(f(x)) = male$ and $f(gender(x)) = male$.

    \subparagraph{(b)}
    For any $x \in P$, we have $clan(m(x)) = clan(x)$ and $m(clan(x)) = m(x)$ (because $m(y) = y$ for both
    $y \in G$). Likewise $clan(f(x))$ is the clan that is not $clan(m(x))$, i.e. the clan that is not $clan(x)$,
    and $f(clan(x))$ is the clan that is not $clan(x)$.

    \subparagraph{(c)}
    \begin{align*}
        f(he-wolf) & = he-bear \\
        f(he-bear) & = he-wolf \\
        f(she-wolf) & = he-bear \\
        f(she-bear) & = he-wolf \\
        m(he-wolf) & = she-wolf \\
        m(he-bear) & = she-bear \\
        m(she-wolf) & = she-wolf \\
        m(she-bear) & = she-bear
    \end{align*}

    \chapter{Session 14. Maps preserve positive properties}

    \paragraph{Exercise 1}
    \[ \beta(y_1) = \beta(f(x_1)) = f(\alpha(x_1)) = f(\alpha(x_2)) = \beta(f(x_2)) = \beta(y_2) \]

    \paragraph{Exercise 2}
    \[ y_2 = f(x_2) = f(\alpha^5(x_1)) = \beta^5(f(x_1)) = \beta^5(y_1) \]

    \paragraph{Exercise 3}
    \[ \beta(y) = \beta(f(x)) = f(\alpha(x)) = f(x) = y \]

    \paragraph{Exercise 4}
    Take $X = \{ 0,1 \}$ with $\alpha(0) = \alpha(1) = 1$ and $Y = \{ * \}$ with $\beta(*) = *$.
    Let $f$ be the only function $X \rightarrow Y$. Then 0 is not a fixed point of $\alpha$
    but $f(0) = *$ is a fixed point of $\beta$.

    \paragraph{Exercise 5}
    If $\alpha^4(x) = x$ then
    \[ y = f(x) = f(\alpha^4(x)) = \beta^4(f(x)) = \beta^4(y) \enspace . \]

    Now, let $X = \{ 0, 1, 2, 3 \}$ with $\alpha(0) = 1$, $\alpha(1) = 2$, $\alpha(2) = 3$ and $\alpha(3) = 0$.
    Let $Y = \{ 0,1 \}$ with $\beta(0) = 1$ and $\beta(1) = 0$. Define $f : X \rightarrow Y$ by $f(0) = 0$,
    $f(1) = 1$, $f(2) = 0$ and $f(3) = 1$. Let $x = 0$ and $y = f(x) = 0$. Then $\alpha^4(x) = x$ but
    $\alpha^2(x) = 2 \neq x$, $\beta^2(y) = y$ but $\beta(y) = 1 \neq y$.

    \chapter{Session 15. Objectification of properties in dynamical systems}
    \section{2. Naming the elements that have a given period by maps}

    \paragraph{Exercise 1}
    Let $x$ be an element in $(X, \alpha)$ of period 5 and 7. Then $\alpha^5(x) = x$ and
    \begin{align*}
        \alpha^7(x) & = x \\
        \therefore \alpha^2(\alpha^5(x)) & = x \\
        \therefore \alpha^2(x) & = x \\
        \alpha^5(x) & = x \\
        \therefore \alpha^3(\alpha^2(x)) & = x \\
        \therefore \alpha^3(x) & = x \\
        \therefore \alpha(\alpha^2(x)) & = x \\
        \therefore \alpha(x) & = x
    \end{align*}

    \paragraph{Exercise 2}
    There are four such maps. For each $x \in C_4$, there is the map $f$ that sends $n$ to $\alpha^n(x)$,
    where $\alpha$ is the endomap on $C_4$.

    \paragraph{Exercise 3}
    Given a map $f : N \rightarrow Y$ in $\mathcal{S}^\circ$, we have
    \begin{align*}
        iteration(evaluation at 0(f))(n) & = iteration(f(0))(n) \\
        & = \beta^{n}(f(0)) \\
        & = f(\sigma^n(0)) \\
        & = f(n) & (\text{for all $n$})\\
        \therefore iteration(evaluation at 0(f)) & = f
    \end{align*}
    Given $y \in Y$, we have
    \begin{align*}
        evaluation at 0(iteration(y)) & = iteration(y)(0) \\
        & = \beta^0(y) \\
        & = y
    \end{align*}

    \paragraph{Exercise 4}
    This holds because $\alpha \circ \alpha = \alpha \circ \alpha$.

    \paragraph{Exercise 5}
    If $evaluation at 0(f) = y$ i.e. $f(0) = y$ then
    \[ evaluation at 0(f \circ \sigma) = f(\sigma(0)) = \beta(f(0)) = \beta(y) \]

    \section{4. The philosophical role of $N$}

    \paragraph{Exercise 6}
    For any person $x$, we have $g(m(x)) = female = m(g(x))$ since $m(y) = female$ for both $y \in B$.

    \paragraph{Exercise 7}
    If we choose $\overline{a} = w$, then there are two choices for $\overline{b}$ (namely $x$ and $z$),
    one choice for $\overline{c}$ (namely $y$), and two choices for $\overline{d}$ (namely $l$ and $m$),
    hence 4 choices in total.

    If we choose $\overline{a} = x$, then there is one choice for $\overline{b}$ (namely $y$), one choice for
    $\overline{c}$ (namely $w$), and two choices for $\overline{d}$ (namely $l$ and $m$),
    hence 2 choices in total.

    If we choose $\overline{a} = y$, then there is one choice for $\overline{b}$ (namely $w$), two choices for
    $\overline{c}$ (namely $x$ and $z$), and two choices for $\overline{d}$ (namely $l$ and $m$),
    hence 4 choices in total.

    If we choose $\overline{a} = z$, then there is one choice for $\overline{b}$ (namely $y$), one choice for
    $\overline{c}$ (namely $w$), and two choices for $\overline{d}$ (namely $l$ and $m$),
    hence 2 choices in total.

    In total, there are 14 maps from $(X, \alpha)$ to $(Y, \beta)$.

    \paragraph{Exercise 10}
    List of labels: $a$, $b$, $c$

    Rules:
    \begin{align*}
        \alpha a & = \alpha b \\
        \alpha^2 a & = \alpha^2 c
    \end{align*}

    \paragraph{Exercise 12}
    Let $f : \mathbb{N} \rightarrow \mathbb{N} \times S$ be the unique map such that $f(0) = (0, s_0)$
    and $f(n+1) = \rho(f(n))$. Let $u = \pi_2 \circ f : \mathbb{N} \rightarrow S$. Prove by induction on $n$
    that $f(n) = (n, u(n))$ for all $n$, hence $u(n+1) = r(n, u(n))$ for all $n$.

    If $v : \mathbb{N} \rightarrow S$ satisfies the same condition, prove $u(n) = v(n)$ for all $n$ by
    induction on $n$.

    \chapter{Session 16. Idempotents, involutions, and graphs}

    \section{2. Solving exercises on maps of graphs}

    \paragraph{Exercise 1}
    They are isomorphic. Let $\mathcal{D}$ be the category defined in Article III Exercise 17.

    Define the functor $F : \mathcal{D} \rightarrow \mathcal{C} / G$ as follows. $F(M, F, \phi, \phi', \mu, \mu')
    = f : (M \cup F, \phi \cup \phi', \mu \cup \mu') \rightarrow G$ where $f(x) = male$ if $x \in M$ and $f(x)
    = female$ if $x \in F$.

    Given $(g, h) : (M, F, \phi, \phi', \mu, \mu') \rightarrow (N, G, \psi, \psi', \nu, \nu')$, let
    $F(g,h) = g \cup h : M \cup F \rightarrow N \cup G$.

    Then $F$ is an isomorphism.

\end{document}